---
title: "codebook.md"
author: "Kisoo Kim"
date: "Oct 2, 2017"
output: html_document
---

# Getting and Cleaning Data Course Project

In this project, you find:

* run_analysis.R : the R-code run on the data set

* Tidy.txt : the clean data extracted from the original data using run_analysis.R

* codeBook.md : the CodeBook 




Getting and Cleaning Data Course Projectless 
The purpose of this project is to demonstrate your ability to collect, work with, and clean a data set. The goal is to prepare tidy data that can be used for later analysis. You will be graded by your peers on a series of yes/no questions related to the project. You will be required to submit: 1) a tidy data set as described below, 2) a link to a Github repository with your script for performing the analysis, and 3) a code book that describes the variables, the data, and any transformations or work that you performed to clean up the data called CodeBook.md. You should also include a README.md in the repo with your scripts. This repo explains how all of the scripts work and how they are connected.

One of the most exciting areas in all of data science right now is wearable computing - see for example this article . Companies like Fitbit, Nike, and Jawbone Up are racing to develop the most advanced algorithms to attract new users. The data linked to from the course website represent data collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at the site where the data was obtained:

http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones

Here are the data for the project:

https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip

You should create one R script called run_analysis.R that does the following.

1. Merges the training and the test sets to create one data set.
2. Extracts only the measurements on the mean and standard deviation for each measurement.
3. Uses descriptive activity names to name the activities in the data set
4. Appropriately labels the data set with descriptive variable names.
5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.


# Description of the Data
The dataset includes the following files:

* 'README.txt'

* 'features_info.txt': Shows information about the variables used on the feature vector.

* 'features.txt': List of all features.

* 'activity_labels.txt': Links the class labels with their activity name.

* 'train/X_train.txt': Training set.

* 'train/y_train.txt': Training labels.

* 'test/X_test.txt': Test set.

* 'test/y_test.txt': Test labels.

* 'train/subject_train.txt': Each row identifies the subject who performed the activity for each window sample. Its range is from 1 to 30. 

* 'train/Inertial Signals/total_acc_x_train.txt': The acceleration signal from the smartphone accelerometer X axis in standard gravity units 'g'. Every row shows a 128 element vector. The same description applies for the 'total_acc_x_train.txt' and 'total_acc_z_train.txt' files for the Y and Z axis. 

* 'train/Inertial Signals/body_acc_x_train.txt': The body acceleration signal obtained by subtracting the gravity from the total acceleration. 

* 'train/Inertial Signals/body_gyro_x_train.txt': The angular velocity vector measured by the gyroscope for each window sample. The units are radians/second. 


The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz. Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (tBodyAcc-XYZ and tGravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz. 

Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (tBodyAccJerk-XYZ and tBodyGyroJerk-XYZ). Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (tBodyAccMag, tGravityAccMag, tBodyAccJerkMag, tBodyGyroMag, tBodyGyroJerkMag). 
Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing fBodyAcc-XYZ, fBodyAccJerk-XYZ, fBodyGyro-XYZ, fBodyAccJerkMag, fBodyGyroMag, fBodyGyroJerkMag. (Note the 'f' to indicate frequency domain signals). 

These signals were used to estimate variables of the feature vector for each pattern:  
'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions.


# Code explanation
## 0. First, load libraries and data.

```{r}
# Load libraries.
library("data.table")

# Load txt files.
dest1 <- "./UCI HAR Dataset/test/X_test.txt"
dest2 <- "./UCI HAR Dataset/test/y_test.txt"
dest3 <- "./UCI HAR Dataset/test/subject_test.txt"
dest4 <- "./UCI HAR Dataset/train/X_train.txt"
dest5 <- "./UCI HAR Dataset/train/y_train.txt"
dest6 <- "./UCI HAR Dataset/train/subject_train.txt"
dest7 <- "./UCI HAR Dataset/features.txt"
dest8 <- "./UCI HAR Dataset/activity_labels.txt"
x_test <- read.table(dest1)             # dim(x_test): 2947 561
y_test <- read.table(dest2)             # dim(y_test): 2947 1
sub_test <- read.table(dest3)           # dim(sub_test): 2947 1
x_train <- read.table(dest4)            # dim(x_train): 7352 561
y_train <- read.table(dest5)            # dim(y_train): 7352 1
sub_train <- read.table(dest6)          # dim(sub_train): 7352 1
feature <- read.table(dest7)            # dim(feature): 561 2
act_labels <- read.table(dest8)         # dim(act_labels): 6 2
```

### 1. Merges the training and the test sets to create one data set.

```{r eval=FALSE}
library("dplyr")
## Add data type {test, train}.
x_test <- mutate(x_test, datatype="test")
x_train <- mutate(x_train, datatype="train")

## Combine each y_test and y_train with act_labels
y_test_labeled <- mutate(y_test, act_label=act_labels[2][V1,])
y_train_labeled <- mutate(y_train, act_label=act_labels[2][V1,])

## Combine data by column
test_combined <- cbind(sub_test, y_test_labeled, x_test)
train_combined <- cbind(sub_train, y_train_labeled, x_train)

## Rename for convenience
names(test_combined)[1] <- "subject"
names(train_combined)[1] <- "subject"
names(test_combined)[2] <- "act_no"
names(train_combined)[2] <- "act_no"

## Then combine the two data by row, and update column name by
## the features
all_combined <- rbind(test_combined, train_combined)
names(all_combined)[4:564] <- as.character(feature[,2])
```

### 2. Extracts only the measurements on the mean and standard deviation for each measurement.

```{r eval=FALSE}
## These are the letters we want to be in the column names
wanted <- c("[Mm]ean\\(\\)","std\\(\\)")
## Select column numbers that contain those letters.
cols_wanted <- 
    c(
        1,2,3,grep(paste(wanted,collapse="|"),
        names(all_combined))
    )
## Then get the subset from the data.
all_wanted <- all_combined[ , cols_wanted]
```



### 3. Uses descriptive activity names to name the activities in the data set

```{r eval=FALSE}
# This was already done by the following codes in
# line 38 and 39.
# y_test_labeled <- mutate(y_test, act_label=act_labels[2][V1,])
# y_train_labeled <- mutate(y_train, act_label=act_labels[2][V1,])
```



### 4. Appropriately labels the data set with descriptive variable names.
```{r eval=FALSE}
names(all_wanted)<-gsub("^t", "time", names(all_wanted))
names(all_wanted)<-gsub("^f", "frequency", names(all_wanted))
names(all_wanted)<-gsub("Acc", "Accelerometer", names(all_wanted))
names(all_wanted)<-gsub("Gyro", "Gyroscope", names(all_wanted))
names(all_wanted)<-gsub("Mag", "Magnitude", names(all_wanted))
names(all_wanted)<-gsub("BodyBody", "Body", names(all_wanted))
```


### 5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.
```{r eval=FALSE}
library(plyr)
head(names(all_wanted))
tidy_result <-aggregate(. ~subject + act_label, all_wanted, mean)
tidy_result<-tidy_result[order(tidy_result$subject,tidy_result$act_no),]
write.table(tidy_result, file = "Tidy.txt",row.name=FALSE)